Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-24 RNARedPrint.tex

\parHead{Specificities and similarities of multi-target design.}
The available methods for multi-target RNA design~\citep{Lyngsoe2012,HoenerzuSiederdissen2013,Taneda2015,Hammer2017} all follow the same overall generate-and-optimize strategy.
%
Faced with the complex constraints due to the multiple targets, early methods such as \Software{Frnakenstein}~\citep{Lyngsoe2012} and \Software{Modena}~\citep{Taneda2015} \revised{do not even attempt to sample sequences systematically from a controlled distribution}, but rely on ad-hoc generation strategies.
%
Recently, the approach \Software{RNAdesign}~\citep{HoenerzuSiederdissen2013}, coupled with local
search in \RNAblueprint{}~\citep{Hammer2017}, solved the
problem of sampling seeds from the \emph{uniform} distribution for multiple
target structures. \Software{RNAdesign} adopts a graph coloring perspective,
coloring the sequence positions with nucleotides, such that the ends of each base pair receive compatible colors. Initially the method decomposes the graph hierarchically, using ear decomposition, and then \Def{precomputes} the number of valid sequences
within each subgraph. The decomposition is then reinterpreted as a
decision tree to perform \Def{stochastic backtracking}, inspired by
Ding and Lawrence~\citep{Ding2003}. Uniform sampling is achieved by
choosing individual nucleotide assignments with probabilities derived
from the subsolution counts. While, due to its decomposition strategy, 
\Software{RNAdesign} performs much better than the theoretical bound of $O(4^n)$, no attempts were made to characterize or justify its---still exponential---complexity; leaving important theoretical questions of the complexity of counting and uniform sampling open.
As well, the \Software{RNAdesign}/\RNAblueprint{} approach is specialized to uniform sampling, which limits its direct extensibility. Both aspects call for a more systematic redesign of multi-target sampling, which builds on established concepts in computer science. In turn, this promises a fundamentally broader range of applications for extensions of the sampling method.

@@

\revised{As central contribution, we provide a systematic and flexibly extensible technique for sampling that targets multiple versatile features. For the sake of clarity, we introduce this method specialized to the sampling of sequences that target the energies of multiple structures and the \GCb\%.}

>>>>

\parHead{Specificities and similarities of multi-target design.}
The available methods for multi-target RNA design~\citep{Lyngsoe2012,HoenerzuSiederdissen2013,Taneda2015,Hammer2017} all follow the same overall generate-and-optimize strategy.
%
Faced with the complex constraints due to the multiple targets, early methods such as \Software{Frnakenstein}~\citep{Lyngsoe2012} and \Software{Modena}~\citep{Taneda2015} \revised{do not even attempt to sample sequences systematically from a controlled distribution}, but rely on ad-hoc generation strategies.
%
Recently, the approach \Software{RNAdesign}~\citep{HoenerzuSiederdissen2013}, coupled with local
search in \RNAblueprint{}~\citep{Hammer2017}, solved the
problem of sampling seeds from the \emph{uniform} distribution for multiple
target structures. \Software{RNAdesign} adopts a graph coloring perspective,
coloring the sequence positions with nucleotides, such that the ends of each base pair receive compatible colors. Initially, the method decomposes the graph hierarchically and then \Def{precomputes} the number of valid sequences
within each subgraph. The decomposition is then reinterpreted as a
decision tree to perform \Def{stochastic backtracking}, inspired by
Ding and Lawrence~\citep{Ding2003}. Uniform sampling is achieved by
choosing individual nucleotide assignments with probabilities derived
from the subsolution counts. While, due to its decomposition strategy, 
\Software{RNAdesign} performs much better than the theoretical bound of $O(4^n)$, no attempts were made to characterize or justify its---still exponential---complexity; leaving important theoretical questions of the complexity of counting and uniform sampling open.
As well, the \Software{RNAdesign}/\RNAblueprint{} approach is specialized to uniform sampling, which limits its direct extensibility. Both aspects call for a more systematic redesign of multi-target sampling, which builds on established concepts in computer science. In turn, this promises a fundamentally broader range of applications for extensions of the sampling method.

@@

\revised{As central contribution, we provide a systematic and flexibly extensible technique for sampling that targets multiple versatile features. For the sake of clarity, we introduce this method specialized to the sampling of RNA sequences that target the energies of multiple structures and the \GCb\%.}
